---
title    : "Class Prediction"
output   : html_document
author   : Akshay Kulkarni
---

<!--
Goal:- To Predict the way in which candidates did their exercise

Sections:-
  * Read and Clean the Data
  * Build The model
  * Train the model
  * Use cross validation to get out of sample error
  * Reasoning behind your choices
  * Use the prediction model to build 20 different cases
---!>

### Get and Clean the Data

Load the training and testing data. Figure out which of columns are not required for building the model. This can be done finding out columns which have all same values or all values are NA i.e values in column are not significantly different.

```{r,eval=FALSE}
library(caret)
library(ggplot2)

set.seed(123)
training  <- read.csv("pml-training.csv",na.strings=c("","NA"))
testing   <- read.csv("pml-testing.csv",na.strings=c("","NA"))

```

Figure out all the columns where sum of all the values in that columns is 0. In this way we can get rid of unnecessary columns

```{r,eval=FALSE}
bad_column_test = lapply(training, function(x) {sum(is.na(x))})
bad_column_name = names(bad_column_test[bad_column_test !=0])

```

On observing the values of first seven columns, we find that they do not contribute anything towards final result and hence can be safely discarded.

```{r,eval=FALSE}
ptraining <- training[,!names(training) %in% bad_column_name]
ptraining <- ptraining[,-c(1:7)]
ptesting  <- testing[,!names(testing) %in% bad_column_name]
ptesting  <- ptesting[,-c(1:7)]
```

### Creating Training and Cross Validation Sets

After cleaning the data, we divide data into training and cross validation set. This is done by using createDataPartition function of caret Package

```{r,eval=FALSE}
inTrain   <- createDataPartition(y=ptraining$classe,p=0.7,list=FALSE)
tr        <- ptraining[inTrain,]
cv        <- ptraining[-inTrain,]
```

After seperating training and cross-validation set, we train the model using Random Forests classifier as they seem to give highest accuracy.

### Build Model
```{r,eval=FALSE}
rfmodel   <- train(tr$classe~.,method="rf",data=tr)
saveRDS(rfmodel, "rfmodel.RDS")
```

Since training and building the model takes huge amount of time, we have stored the model on file and can be later used without actually again creating it.

### Error estimation with cross validation
The model is now used to perform a cross validation with cross-validation data set.
The out of error rate is expected to be less than 1%, as the accuracy of the model observed above is 99.88%.

```{r,eval=FALSE}
mean(predict(rfmodel, ptesting) == ptesting$classe) * 100

tc        <- trainControl(method="cv",number=3)
rfcvmodel <- train(cv$classe~.,method="rf",data=cv,trControl=tc)
saveRDS(rfcvmodel,"rfcvmodel.RDS")
predcv    <- predict(rfcvmodel,cv)

confusionMatrix(predcv,cv$classe)
```

### Predict the 20 test cases
To predict the classe of the testing dataset, we're applying the prediction using the model we've trained.

```{r,eval=FALSE}
test_prediction<-predict(rfmodel, newdata=ptesting)
test_prediction
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(test_prediction)
```
